---
title: "Bank Marketing"
author: "Joshua Williams, Raymond Obidegwu, Heath McCasland"
date: "July 1, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Import dependencies
```{r}
library(ResourceSelection)
library(h2o)
```


## Read in the data 
The data was donated in 2012 to UCI by a portuguese bank. It in most likely BPI given the size of the company. 
```{r}
bank = read.table("C:/Users/willi/Desktop/bank-additional/bank-additional/bank-additional-full.csv", sep=";",header=T)

```

Input variables:
###bank client data:
1 - age (numeric)
2 - job : type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')
3 - marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
4 - education (categorical: 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')
5 - default: has credit in default? (categorical: 'no','yes','unknown')
6 - housing: has housing loan? (categorical: 'no','yes','unknown')
7 - loan: has personal loan? (categorical: 'no','yes','unknown')
###related with the last contact of the current campaign:
8 - contact: contact communication type (categorical: 'cellular','telephone') 
9 - month: last contact month of year (categorical: 'jan', 'feb', 'mar', ..., 'nov', 'dec')
10 - day_of_week: last contact day of the week (categorical: 'mon','tue','wed','thu','fri')
11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
###other attributes:
12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
14 - previous: number of contacts performed before this campaign and for this client (numeric)
15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')
###social and economic context attributes
16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
17 - cons.price.idx: consumer price index - monthly indicator (numeric) 
18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric) 
19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
20 - nr.employed: number of employees - quarterly indicator (numeric)

Output variable (desired target):
21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

```{r}
str(bank)

```

###Data Cleaning
As mentioned above the duration is a very good predictor, but not important considering our business question. It is intuitive that if a call only last a few seconds there was probably no sale made. If a call last longer a sale is more likely to have been made, becuase more time is available to awnser the customers concerns before they invest. 
```{r}
table(bank[which(bank$duration < 38),]$y)
```

There would be no way to know the duration of the call before the call is made. The calls made with less than 4 seconds are deleted. 
```{r}
rows=row.names(bank[which(bank$duration < 4),])
rows=as.integer(rows)
bank = bank[-rows,]
dim(bank)

```

Duration is not used becuase it is not known how long a phone call will last before it is made. The month and day of the week columns are also deleted, because they are irelevant to the business question on hand. 

```{r}
bank$duration = NULL
bank$month=NULL
bank$day_of_week=NULL
```

Normalize the data.
```{r}
norm= function(x){
  return((x-min(x))/ (max(x)- min(x)))}

for (i in 1:dim(bank)[2]){ bank[,i]=as.numeric(bank[,i])}

banknorm= data.frame((lapply(bank[,1:(dim(bank)[2]-1)], norm)))
banknorm$y= bank$y
```


###Data Exploratoion

It appears there is some correlation of between the social economic factors. 

```{r}
corrplot::corrplot(stats::cor(banknorm),method="number")
```

Bank Client Data and Predictor 
```{r}
ResourceSelection::kdepairs(banknorm[,c(1,2,3,4,6,18)])
stats::cor(banknorm)[c(5,7),]
```
Correlation between 
```{r}
corrplot::corrplot(stats::cor(banknorm[7:12]),method="number")
```



There appears to be some multicollinearity between the social and economic factors. 
```{r}
ResourceSelection::kdepairs(banknorm[13:18])
```


```{r}

pcse= princomp(banknorm[13:17])
summary(pcse)
Socio_Economic = pcse$scores[,1:2] #92% cumulative variennce
#remove old socio economic 
banknorm[13:17]=NULL
banknorm$SEC1=Socio_Economic[,1]
banknorm$SEC2=Socio_Economic[,2]
```

The correlation plot shows that the priciple components fixed the multicollinearity. 
```{r}
corrplot::corrplot(stats::cor(banknorm),method="number")
```

Return y to factor so Neural net can output probability with softmax.
```{r}
banknorm$y=as.factor(banknorm$y)
```


Split the data into train and test.

```{r}
idx = sample(1:nrow(banknorm), 0.7*nrow(banknorm))

train=banknorm[idx,]
test=banknorm[-idx,]
dim(test)
dim(train)

```



```{r}
h2o.init()
x.hex = as.h2o(train)
test.hex = as.h2o(test)

deep = h2o.deeplearning(training_frame=x.hex,
                 x=c(1:12,14,15),
                 y=13,
                 epochs = 10,
                 activation="Rectifier",
                 hidden=c(10,10),
                 rate = .03,
                 adaptive_rate = T,
                 nesterov_accelerated_gradient = T,
                 initial_weight_distribution = "UniformAdaptive")
```

Predict the test data. 
```{r}
pred= predict(deep, test.hex)

```

The AUC Curve is plotted below. As seen the model has a decent performance. 

```{r}
perf = h2o.performance(deep, test.hex)
plot(perf)
```

The model has good performance, the model guessed that 
```{r}
perf
```

The original data will be used with the same index as test in order to see the data without normilazation.It appears the model ouput
The people most likely to buy are:
age: 42
job: Admin, then Blue collar
Marital: Married
Education: University Degree
Default: NO
loan : NO
conatct: Cellular
Number of contact performed: 1
Number of days passed before client was previously contacted: 999
poutcome: nonexistent



```{r}
testdenorm=bank[-idx,] 
testdenorm$pred=as.data.frame(pred)$predict
test$pred=as.data.frame(pred)$predict
yes=dplyr::filter(testdenorm, pred==2)
mean(yes$age)
hist(yes$age)

```



```{r}
nb = h2o.naiveBayes(training_frame=x.hex,
                 x=c(1:12,14,15),
                 y=13)

```


```{r}
perfnb = h2o.performance(nb, test.hex)
plot(perfnb)
perfnb
```


```{r}
percent_yes = table(bank$y)[[2]]/dim(bank)[1]
print(paste("percent of people who say yes", percent_yes))
bank = read.table("C:/Users/willi/Desktop/bank-additional/bank-additional/bank-additional-full.csv", sep=";",header=T)
mean_duration = (mean(bank$duration)/60)+2 
print(paste("Avg Duration of call in min plus two minutes for transition:" ,mean_duration))
hrl_rate=31.5 #average hrly salary of bank sales man in US
num_cust=1400000 #The number of employees was aproximated by multipling 5191(from) data and 280 (customer to employee ratio, retreived from competitors)
```

In order to know how many customers to call a monte carlo simulation is run to capture customers that say yes from the .11 percent of the customers that actually say yes. 
```{r}
percent_called=1:100
people =c() #pectent of the 11% of customers who will buy
for(q in 1:100){
#Monte Carlo Simulation
l=c()
for (i in 1:250){
samp = sample(1:(dim(test)[1]), 1000)
df=test[samp,]
table(test$y)[[2]]/table(test$y)[[1]]
df1=df[order(df$pred,decreasing = T),]
dfcall= df1[1:(1000*q/100),]
l=append(l,(table(dfcall$y)[[2]]/table(df$y)[[2]]))
}
people=append(people, mean(l))
}
da=data.frame(percent_called, people)
plot(da)


#Same plot but with number of people on the yaxis rather than percent of percent. 
percent_called=1:100
people =c() #pectent of the 11% of customers who will buy
for(q in 1:100){
#Monte Carlo Simulation
l=c()
for (i in 1:250){
samp = sample(1:(dim(test)[1]), 1000)
df=test[samp,]
table(test$y)[[2]]/table(test$y)[[1]]
df1=df[order(df$pred,decreasing = T),]
dfcall= df1[1:(1000*q/100),]
l=append(l,(table(dfcall$y)[[2]]/table(df$y)[[2]]))
}
people=append(people, ((mean(l)*0.1126542)*num_cust) )
}
da=data.frame(percent_called, people)
plot(da)

```

Final Simulation using 18 percent of customers.
```{r}
#Monte Carlo Simulation
l=c()
for (i in 1:1000){
samp = sample(1:(dim(test)[1]), 1000)
df=test[samp,]
table(test$y)[[2]]/table(test$y)[[1]]
df1=df[order(df$pred,decreasing = T),]
dfcall= df1[1:(1000*.18),]
l=append(l,(table(dfcall$y)[[2]]/table(df$y)[[2]]))
}


#cost without model
WM_cost_call=(num_cust/mean_duration)*hrl_rate
  
#cost with model
M_cost_call=((num_cust*.18)/mean_duration)*hrl_rate

#Savings
s=WM_cost_call-M_cost_call
s=formatC(s, format="d", big.mark=",")
lower = formatC(as.integer(((mean(l)-2*sd(l)))*(percent_yes*num_cust)), format="d", big.mark=",")
upper=formatC(as.integer((mean(l)+2*sd(l))*(percent_yes*num_cust)), format="d", big.mark=",")

#Certainty from monte carlo simulation
certainty= sum(l>(mean(l)-2*sd(l)) & l<(mean(l)+2*sd(l)))/length(l)
hist((l*(percent_yes*num_cust)), main="Distribution of the Results of Simulation", xlab= "Number of Customers")
print(paste("Based on the simulation by calling 18% of the customers there is ",certainty*100,"%"," certainty that the model will get between ",lower," and " ,upper," customers saving $",s, sep=''))

```











